Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-04-15 main.tex
Fixed Wegman citation key.
Use verbatim for the github url because the url command is not available when biblatex is not used.

\section{Introduction}
%
Single-actor bicycle crashes are associated with a surprisingly large
percentage of reported serious injuries~\citep{Wegman2024a}. At low speeds,
from start up to typical cruising speeds, bicycles are not self-stable and can
be challenging for the rider to balance. Low speed crashes may be reduced if
the bicycle was self-stable at these speeds by relieving the rider of some of
their necessary control activity. Bicycles can be mechanically modified to
lower the speeds at which they are self-stable~\citep{Astrom2005} and since the
1980s it has been known that applying a motor actuated steering torque
proportional to the vehicle's roll angular rate~\footnote{Roll angular rate is
defined herein as per the SAE J670 vehicle standards.} can stabilize a
single-track vehicle down to very low speeds~\citep{Ruijs1986}. If automatic
control of steering can stabilize a bicycle, it may reduce the control required
from the rider to successfully manage balancing tasks much like natural high
speed self-stability already does. We have developed a balance-assisting
bicycle, Figure~\ref{fig:balance-assist-bicycle}, based on these principles and
hypothesize that it helps the rider in situations in which they are likely to
fall.
%
\begin{figure}
  \centering
  \includegraphics[width=80mm]{figures/balance-assist-bicycle.jpg}
  \caption{Balance assist bicycle prototype with electric motor in the steering
    column and data acquisition and control electronics mounted in the rear
    rack first introduced in \citet{Alizadehsaravi2023}.}
  \label{fig:balance-assist-bicycle}
\end{figure}

@@

\noindent
\url{https://github.com/mechmotum/fall-probability-paper}

>>>>

\section{Introduction}
%
Single-actor bicycle crashes are associated with a surprisingly large
percentage of reported serious injuries~\citep{Wegman2024}. At low speeds,
from start up to typical cruising speeds, bicycles are not self-stable and can
be challenging for the rider to balance. Low speed crashes may be reduced if
the bicycle was self-stable at these speeds by relieving the rider of some of
their necessary control activity. Bicycles can be mechanically modified to
lower the speeds at which they are self-stable~\citep{Astrom2005} and since the
1980s it has been known that applying a motor actuated steering torque
proportional to the vehicle's roll angular rate~\footnote{Roll angular rate is
defined herein as per the SAE J670 vehicle standards.} can stabilize a
single-track vehicle down to very low speeds~\citep{Ruijs1986}. If automatic
control of steering can stabilize a bicycle, it may reduce the control required
from the rider to successfully manage balancing tasks much like natural high
speed self-stability already does. We have developed a balance-assisting
bicycle, Figure~\ref{fig:balance-assist-bicycle}, based on these principles and
hypothesize that it helps the rider in situations in which they are likely to
fall.
%
\begin{figure}
  \centering
  \includegraphics[width=80mm]{figures/balance-assist-bicycle.jpg}
  \caption{Balance assist bicycle prototype with electric motor in the steering
    column and data acquisition and control electronics mounted in the rear
    rack first introduced in \citet{Alizadehsaravi2023}.}
  \label{fig:balance-assist-bicycle}
\end{figure}

@@

\noindent
\verb|https://github.com/mechmotum/fall-probability-paper|
